---
title: "RWorksheet_Gerona#4c"
author: "Mariel M. Gerona"
output: pdf_document
---

#1a: Importing and Exploring the mpg Dataset
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# Load the mpg dataset
data("mpg")

# Display first few rows
head(mpg)
```

#1b. Which variables are categorical?
```{r}
categorical_vars <- mpg %>% select(where(is.character))
categorical_vars

```
#1c. Which variables are continuous?
```{r}
continuous_vars <- mpg %>% select(where(is.numeric))
continuous_vars

```


#2a. Group Manufacturers and Unique Models
```{r}
manufacturer_models <- mpg %>% 
  group_by(manufacturer) %>% 
  summarise(unique_models = list(unique(model)))

manufacturer_models

```
#2. Plot Manufacturer Model Counts
```{r}
# Count the models per manufacturer
manufacturer_counts <- mpg %>%
  group_by(manufacturer) %>%
  summarise(models = n_distinct(model))

# Plot the counts of models per manufacturer
ggplot(manufacturer_counts, aes(x = reorder(manufacturer, -models), y = models)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(title = "Number of Models per Manufacturer", x = "Manufacturer", y = "Model Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
##2b. Graph Unique Models Using plot() and ggplot()
```{r}
sorted_data <- manufacturer_counts[order(manufacturer_counts$models, decreasing = TRUE),]
barplot(sorted_data$models, names.arg = sorted_data$manufacturer, las = 2, col = "Pink",
main = "Number of Unique Models by Manufacturer",
xlab = "Manufacturer", ylab = "Number of Unique Models")

```

```{r}
# Using ggplot2
ggplot(manufacturer_counts, aes(x = reorder(manufacturer, -models), y = models)) +
  geom_bar(stat = "identity", fill = "coral") +
  labs(title = "Unique Models per Manufacturer (ggplot2)",
       x = "Manufacturer",
       y = "Number of Unique Models") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#2. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
ggplot(mpg, aes(x = model, y = manufacturer)) + geom_point()

```


#2a. This plot shows the relationship between model and manufacturer. However, it may not be very informative due to the large number of model names, which can overlap and create a cluttered appearance, making the plot difficult to interpret.

#3. Plot the model and the year using ggplot(). Use only the top 20 observations.
```{r}
# Get the top 20 observations based on year
top_20_mpg <- mpg %>% 
  arrange(desc(year)) %>% 
  head(20)

# Create the plot
ggplot(top_20_mpg, aes(x = model, y = year)) +
  geom_point(color = "blue") +
  labs(title = "Top 20 Models by Year",
       x = "Model",
       y = "Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
#4a. Plot using geom_bar() using the top 20 observations only. The graphs should have a title, labels, and colors.
```{r}
# Group the data by model and count the number of cars per model
model_counts <- mpg %>%
  group_by(model) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(20)

# Plot the top 20 models with geom_bar()
ggplot(model_counts, aes(x = reorder(model, -count), y = count, fill = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 20 Car Models by Count",
       x = "Car Model",
       y = "Number of Cars",
       fill = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient(low = "lightblue", high = "blue")

```
#4b. Plot using the geom_bar() + coord_flip().
```{r}
# Plot the top 20 models with geom_bar() and coord_flip()
ggplot(model_counts, aes(x = reorder(model, -count), y = count, fill = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 20 Car Models by Count",
       x = "Car Model",
       y = "Number of Cars",
       fill = "Count") +
  theme_minimal() +
  scale_fill_gradient(low = "lightblue", high = "blue")

```
#5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.
```{r}
# Create the plot
ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement (L)") +
  theme_minimal() +
  scale_color_gradient(low = "lightblue", high = "blue")

```

#6. Plot the relationship between displ (engine displacement) and hwy (highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c.
```{r}
# Create the plot
ggplot(mpg, aes(x = displ, y = hwy, color = cyl)) +
  geom_point() +
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement (L)",
       y = "Highway Miles per Gallon",
       color = "Number of Cylinders") +
  theme_minimal()

```

#6a. How many numbers of observation does it have? What are the variables of the traffic dataset?
```{r}
# Load necessary library for reading Excel files
library(readxl)

# Import the traffic.xlsx file
traffic_data <- read_excel("/cloud/project/Worksheet#4/Worksheet#4c/traffic.xlsx")

# Check the number of observations and variables
n_obs <- nrow(traffic_data)  # Number of observations
variables <- colnames(traffic_data)  # Variables in the dataset

# Show results
n_obs
variables

```

#6b. Subset the traffic dataset into junctions.
```{r}
# Subset the dataset based on the identified column (e.g., location)
junctions_data <- traffic_data %>% 
  filter(!is.na(location))  # Replace 'location' with the actual column name

# View the first few rows of the junctions dataset
head(junctions_data)

```


#6c. Plot each junction using geom_line.
```{r}
# Reshape data to long format using pivot_longer
traffic_data_long <- traffic_data %>%
  pivot_longer(cols = starts_with("Junction"),  # Select columns starting with "Junction"
               names_to = "Junction",            # New column to hold the junction names
               values_to = "Traffic")           # New column to hold the traffic data

# Plotting the traffic data using geom_line
ggplot(traffic_data_long, aes(x = Date, y = Traffic, color = Junction, group = Junction)) +
  geom_line() +
  labs(title = "Traffic Trends by Junction",
       x = "Date",
       y = "Traffic Volume") +
  theme_minimal()

```

#7. From alexa_file.xlsx, import it to your environment.
```{r}
# Load necessary libraries
library(readxl)

# Set the correct file path
file_path <- "/cloud/project/Worksheet#4/Worksheet#4c/alexa_file.xlsx"

# Import the alexa_file.xlsx
alexa_file <- read_excel(file_path)

```

#7a. How many observations does alexa_file has? What about the number of columns?
```{r}
# Check the number of observations and columns
num_observations <- nrow(alexa_file)
num_columns <- ncol(alexa_file)

# Show results
num_observations
num_columns

```

#7b. Group the variations and get the total of each variation using dplyr.
```{r}
# Group by variations and get the total of each variation
variation_totals <- alexa_file %>%
  group_by(variation) %>%
  summarise(total = sum(value, na.rm = TRUE))

# Show the results
variation_totals

```

#7c. Plot the variations using ggplot2.
```{r}
ggplot(variation_totals, aes(x = variation, y = total, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Total by Variation", 
       x = "Variation", 
       y = "Total") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#7d. Plot a geom_line() with the date and the number of verified reviews.
```{r}
ggplot(alexa_file, aes(x = date, y = verified_reviews)) +
  geom_line(color = "blue") +
  labs(title = "Verified Reviews Over Time", 
       x = "Date", 
       y = "Number of Verified Reviews") +
  theme_minimal()

```

#7e. Get the relationship of variations and ratings.
```{r}
ggplot(alexa_file, aes(x = variation, y = rating, color = variation)) +
  geom_boxplot() +
  labs(title = "Variation vs. Rating", 
       x = "Variation", 
       y = "Rating") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```






