---
title: "RWorksheet_Gerona#4c"
author: "Mariel M. Gerona"
date: "2024-11-10"
output: pdf_document
---

```{r}
library(ggplot2)
library(dplyr)

# 1. Load the mpg dataset
mpg <- ggplot2::mpg
```


# a. Solution to import a CSV file into the environment (assuming 'mpg.csv' as file name)
)
```{r}
mpg <- read.csv("path/to/mpg.csv"
```

# b. Identifying categorical variables in the mpg dataset
```{r}
categorical_vars <- names(mpg)[sapply(mpg, is.factor)]
print("Categorical variables:")
print(categorical_vars)
```


# c. Identifying continuous variables in the mpg dataset
```{r}
continuous_vars <- names(mpg)[sapply(mpg, is.numeric)]
print("Continuous variables:")
print(continuous_vars)
```

# 2. Manufacturer with the most models and model with the most variations
```{r}
most_models <- mpg %>% 
  group_by(manufacturer) %>% 
  summarize(model_count = n_distinct(model)) %>% 
  arrange(desc(model_count))

most_variations <- mpg %>%
  group_by(model) %>%
  summarize(variation_count = n()) %>%
  arrange(desc(variation_count))
```

# a. Group manufacturers and find unique models
```{r}
unique_models <- mpg %>%
  group_by(manufacturer) %>%
  summarize(unique_models = n_distinct(model))
```

# Plot using base plot
```{r}
plot(unique_models$manufacturer, unique_models$unique_models,
     main="Number of Unique Models per Manufacturer",
     xlab="Manufacturer", ylab="Unique Models", las=2)
```

# Plot using ggplot
ggplot(unique_models, aes(x=manufacturer, y=unique_models)) +
  geom_bar(stat="identity") +
  labs(title="Number of Unique Models per Manufacturer", x="Manufacturer", y="Unique Models") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# 3. Relationship between model and manufacturer using ggplot
ggplot(mpg, aes(x=model, y=manufacturer)) +
  geom_point() +
  labs(title="Scatter plot of Model and Manufacturer")

# 4. Plot model and year for top 20 observations
top_20_mpg <- mpg %>% slice(1:20)
ggplot(top_20_mpg, aes(x=model, y=year)) +
  geom_point() +
  labs(title="Top 20 Models by Year", x="Model", y="Year")

# 5. Group by model and count number of cars per model
car_count <- mpg %>%
  group_by(model) %>%
  summarize(car_count = n()) %>%
  arrange(desc(car_count)) %>%
  slice(1:20)

# a. Plot using geom_bar()
ggplot(car_count, aes(x=reorder(model, -car_count), y=car_count)) +
  geom_bar(stat="identity", fill="blue") +
  labs(title="Top 20 Car Models", x="Model", y="Number of Cars") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# b. Plot using coord_flip
ggplot(car_count, aes(x=reorder(model, car_count), y=car_count)) +
  geom_bar(stat="identity", fill="blue") +
  labs(title="Top 20 Car Models (Flipped)", x="Model", y="Number of Cars") +
  coord_flip()

# 6. Relationship between cylinders and engine displacement
ggplot(mpg, aes(x=cyl, y=displ, color=displ)) +
  geom_point() +
  labs(title="Relationship between No. of Cylinders and Engine Displacement", x="Cylinders", y="Engine Displacement")

# 7. Relationship between engine displacement and highway mpg
ggplot(mpg, aes(x=displ, y=hwy, color=cty)) +
  geom_point() +
  labs(title="Relationship between Engine Displacement and Highway MPG", x="Engine Displacement", y="Highway MPG")

# 8. Importing traffic.csv
# traffic <- read.csv("traffic.csv")

# a. Number of observations and variables in traffic dataset
# traffic_summary <- list(observations = nrow(traffic), variables = names(traffic))

# b. Subset traffic dataset into junctions
# junctions <- traffic %>% group_by(junction) %>% summarize()

# c. Plot each junction using geom_line()
# ggplot(traffic, aes(x=timestamp, y=count, color=junction)) +
#   geom_line() +
#   labs(title="Traffic Counts by Junction", x="Timestamp", y="Count")

# 9. Importing alexa_file.xlsx
# library(readxl)
# alexa_file <- read_excel("alexa_file.xlsx")

# a. Summary of alexa_file
# alexa_summary <- list(observations = nrow(alexa_file), columns = ncol(alexa_file))

# b. Group variations and count each using dplyr
# variation_counts <- alexa_file %>% group_by(variation) %>% summarize(count = n())

# c. Plot variations using ggplot
# ggplot(variation_counts, aes(x=variation, y=count)) +
#   geom_bar(stat="identity", fill="green") +
#   labs(title="Counts of Variations", x="Variation", y="Count")

# d. Plot date and verified reviews with geom_line()
# ggplot(alexa_file, aes(x=date, y=verified_reviews)) +
#   geom_line() +
#   labs(title="Verified Reviews over Time", x="Date", y="Verified Reviews")

# e. Relationship between variations and ratings
# ggplot(alexa_file, aes(x=variation, y=rating)) +
#   geom_boxplot() +
#   labs(title="Ratings by Variation", x="Variation", y="Rating")
