---
title: "RWorksheet_Gerona#3a"
author: "Mariel M. Gerona"
output: pdf_document
---

# 1. There is a built-in vector LETTERS contains the uppercase letters of the alphabet and letters which contains lowercase letters of the alphabet.

# a. Produce a vector that contains the first 11 letters.
```{r}
first_11 <- LETTERS[1:11]
print(first_11)
```


# b. Produce a vector that contains the odd numbered letters.
```{r}
odd_letters <- LETTERS[seq(1, 26, by=2)]
print(odd_letters)

```

# c. Produce a vector that contains the vowel letters.
```{r}
vowel_letters <- LETTERS[c(1, 5, 9, 15, 21)]
print(vowel_letters)
```

 
# d. Produce a vector that contains the last 5 lowercase letters.
```{r}
last_5 <- letters[22:26]
print(last_5)
```

 
# e. Produce a vector that contains letters between 15 to 24 letters in lowercase.
```{r}
letters_between <- letters[seq(15, 24)]
print(letters_between)
```



# 2. Create a vector(not a dataframe) with the average temperatures in April

# a. Character name of cities/towns
```{r}
city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
print(city)
```


# b - Average temperatures in Celcius
```{r}
temp <- c(42, 39, 34, 34, 30, 27)
print(temp)
```


# c - Create a data frame to combine the city and temp
```{r}
city_temp_df <- data.frame(City = city, Temperature = temp)
print(city_temp_df)

```

# d - Rename the columns using names() function as City and Temperature 
```{r}
names(city_temp_df) <- c("City", "Temperature")
print(city_temp_df)

```


# e - Print structure of dataframe by using str() function.
```{r}
str(city_temp_df)
```

# f - Display the content of row 3 and row 4
```{r}
print(city_temp_df[3:4,])
```

# g - Display city with highest and lowest temperature
```{r}
max_temp_city <- city_temp_df$City[which.max(city_temp_df$Temperature)]
min_temp_city <- city_temp_df$City[which.min(city_temp_df$Temperature)]

print(max_temp_city)

print(min_temp_city)

```


# 2. Create a matrix of one to eight and eleven to fourteen with four columns and three rows. 

# a. Create matrix
```{r}
matrix_data <- matrix(c(1:8, 11:14), nrow=3, ncol=4)
print(matrix_data)
```



# b. Multiply the matrix by two.
```{r}
mult_matrix <- matrix_data * 2
print(mult_matrix)
```


# c. Content of row 2
```{r}
row_2 <- matrix_data[2,]
print(row_2)
```


# d. Display output of the columns in 2 and 3, row 3
```{r}
subset_matrix <- matrix_data[2:3, 3:3]
print(subset_matrix)
```


# 3. An array contains 1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1

# 3a - Create an array for the numeric values
```{r}
array_data <- array(c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1), dim = c(2, 4, 3))
print(array_data)
```



# 3b - Display array dimensions
```{r}
dim(array_data)
```


# 3c - Name rows, columns, and dimensions
```{r}
dimnames(array_data) <- list(
  letters[1:2], LETTERS[1:4], c("1st-Dimensional Array", "2nd-Dimensional Array", "3rd-Dimensional Array"))
print(array_data)
```

