---
title: "RWorksheet_Gerona#3b"
output: pdf_document
author: Mariel M. Gerona
---

# 1. Create a data frame using the table below

# a. Create a data frame
```{r}
df <- data.frame(
  Name = c("John", "Alice", "Peter", "Emma", "Mark"),
  Siblings = c(4, 5, 3, 6, 5),
  Sex = c("Male", "Female", "Male", "Female", "Male"),
  Father_Occupation = c("Farmer", "Driver", "Farmer", "Others", "Farmer"),
  Type_of_House = c("Wood", "Concrete", "Semi-Concrete", "Wood", "Concrete"))
```


# b. Get structure and summary of the data
```{r}
str(df)
summary(df)
```


# c. Check if the mean number of siblings is 5
```{r}
mean_siblings <- mean(df$Siblings)
mean_siblings == 5
```


# d. Extract the first two rows and all columns using the subsetting functions
```{r}
first_two_rows <- df[1:2, ]
print(first_two_rows)
```


# e. Extract 3rd and 5th rows with 2nd and 4th columns
```{r}
selected_rows_cols <- df[c(3, 5), c(2, 4)]
print(selected_rows_cols)
```


# f. Select the variable "Type of Houses" then store the vector that results as types_houses
```{r}
types_houses <- df$Type_of_House

```

# g. Select only all male respondents whose father's occupation was farmer
```{r}
males_farmers <- df[df$Sex == "Male" & df$Father_Occupation == "Farmer", ]
print(males_farmers)
```


# h. Select only all female respondents that have greater than or equal to 5 number of siblings attending schools
```{r}
females_siblings <- df[df$Sex == "Female" & df$Siblings >= 5, ]
print(females_siblings)
```


# 2. Create an empty data frame and describe the result
```{r}
df_empty <- data.frame(
  Ints = integer(),
  Doubles = double(),
  Characters = character(),
  Logicals = logical(),
  Factors = factor(),
  stringsAsFactors = FALSE)
```


# Display the structure of the empty dataframe
```{r}
print("Structure of the empty dataframe:")
str(df_empty)
```


# 3. Save the data frame to CSV
```{r}
write.csv(df, "HouseholdData.csv", row.names = FALSE)
```


# a. Import the CSV file into the R environment
```{r}
df_import <- read.csv("HouseholdData.csv")

```

# b. Convert the Sex into factor using factor() function and change it into integer.
```{r}
df_import$Sex <- factor(df_import$Sex, levels = c("Male", "Female"), labels = c(1, 2))
print(df_import)
```


# c. Convert the Type of Houses into factor and change it into integer.
```{r}
df <- data.frame(Type_of_Houses = c("Wood", "Concrete", "Semi-Concretes", "Wood", "Concrete"))
df$Type_of_Houses <- factor(df$Type_of_Houses, levels = c("Wood", "Concrete", "Semi-Concretes"), labels = c(1, 2, 3))
df$Type_of_Houses <- as.integer(df$Type_of_Houses)
print(df)
```


# d. Convert "Father_Occupation" to factor with Farmer = 1, Driver = 2, Others = 3
```{r}
df_import$Father_Occupation <- factor(df_import$Father_Occupation, levels = c("Farmer", "Driver", "Others"), labels = c(1, 2, 3))
print(df_import)
```


# e. Select all Female respondents whose father is a Driver
```{r}
female_driver <- subset(df_import, Sex == 2 & Father_Occupation == 2)
print(female_driver)
```


# f. Select respondents with 5 or more siblings attending school
```{r}
siblings_5_or_more <- subset(df_import, Siblings >= 5)
print(siblings_5_or_more)
```


# 4. Interpret the graph
```{r}
df <- data.frame(
  Name = c("Alice", "Bob", "Charlie", "David"),
  Siblings = c(2, 3, 1, 4))
  
barplot(df$Siblings, names.arg = df$Name, 
        main = "Number of Siblings per Respondent", 
        xlab = "Name", ylab = "Siblings", 
        col = "blue", las = 2)
```

        

