---
title: "RWorksheet#5a_group(Espadon, Bernasol, Gerona)"
output: pdf_document
date: "2024-11-11"
---
#1. Scraping top 50 tv shows in Imdb.com.
```{r}
library(rvest)
library(writexl)

# URL of IMDb's Top TV Shows
url <- "https://www.imdb.com/chart/toptv/"

# Read the page HTML
page <- read_html(url)

# Extract data
ranks <- page %>% html_nodes(".titleColumn") %>% html_text(trim = TRUE) %>% gsub("^[ \t]*([0-9]+).*$", "\\1", .)
titles <- page %>% html_nodes(".titleColumn a") %>% html_text()
years <- page %>% html_nodes(".titleColumn span.secondaryInfo") %>% html_text() %>% gsub("[()]", "", .)
ratings <- page %>% html_nodes(".imdbRating strong") %>% html_text() %>% as.numeric()

# Combine data into a dataframe
imdb_data <- data.frame(Rank = as.integer(ranks), Title = titles, Year = as.integer(years), Rating = ratings)

# Write to Excel file
write_xlsx(imdb_data, "IMDb_Top_TV_Shows.xlsx")

```

#2. Scraping User Reviews for Selected TV Shows
```{r}
get_reviews <- function(show_url) {
  review_page <- read_html(show_url)

  # Extract review details, handling cases where data might be missing
  reviewer_names <- review_page %>%
    html_nodes(".display-name-link a") %>%
    html_text()
  
  review_dates <- review_page %>%
    html_nodes(".review-date") %>%
    html_text()
  
  ratings <- review_page %>%
    html_nodes(".rating-other-user-rating span:nth-child(1)") %>%
    html_text() %>%
    as.numeric()
  
  review_texts <- review_page %>%
    html_nodes(".text.show-more__control") %>%
    html_text()
  
  # Determine the maximum length among the vectors to ensure consistent rows
  max_length <- max(length(reviewer_names), length(review_dates), length(ratings), length(review_texts))
  
  # Pad vectors with NA to ensure they all have the same length
  reviewer_names <- c(reviewer_names, rep(NA, max_length - length(reviewer_names)))
  review_dates <- c(review_dates, rep(NA, max_length - length(review_dates)))
  ratings <- c(ratings, rep(NA, max_length - length(ratings)))
  review_texts <- c(review_texts, rep(NA, max_length - length(review_texts)))
  
  # Combine the data into a data frame
  data.frame(Reviewer = reviewer_names, Date = review_dates, Rating = ratings, Review = review_texts)
}

# Test with the URL
show_url <- "https://www.imdb.com/title/tt0903747/?ref_=chttvtp_t_1"  # Breaking Bad review page
reviews_df <- get_reviews(show_url)

```
#3. Creating a Time Series Graph of TV Shows by Year:
```{r}
library(ggplot2)

ggplot(top_tv_shows, aes(x = Year)) +
  geom_bar() +
  ggtitle("Number of TV Shows Released per Year") +
  xlab("Year") +
  ylab("Count of TV Shows")

```

#4. Selecting Categories and Products:
#5. Scraping Product Details (Price, Description, Ratings, Reviews):
```{r}
get_product_data <- function(product_url) {
  product_page <- read_html(product_url)

  price <- product_page %>%
    html_node(".priceblock_ourprice") %>%
    html_text()

  description <- product_page %>%
    html_node("#productDescription") %>%
    html_text()

  rating <- product_page %>%
    html_node(".a-icon-alt") %>%
    html_text()

  data.frame(Price = price, Description = description, Rating = rating)
}

# Loop over product URLs for a category
product_urls <- c("url1", "url2", "url3")
product_data <- lapply(product_urls, get_product_data)

```

#6. Describing the Extracted Data:

#7. Use case for the data you have extracted

#8. ggplot(product_data, aes(x = Rating, y = Price)) +
```{r}
 geom_point() +
  ggtitle("Price vs. Ratings for Products in Each Category") +
  xlab("Ratings") +
  ylab("Price")

```

#9. Ranking Products by Price and Ratings:
```{r}
product_data_sorted <- product_data[order(product_data$Price, decreasing = TRUE), ]
head(product_data_sorted)
 
```


