---
title: "RWorksheet_Gerona#4"
author: "Mariel M. Gerona"
output: pdf_document
---

# 1a. Describe the Data
```{r}
data <- data.frame(
Shoe_Size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.75, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
)
```



# 1b. Filter females
```{r}
females <- subset(data, Gender == "F", select = c(Shoe_Size, Height))
print("Female data:")
print(females)
```



# 1c. Filter males
```{r}
males <- subset(data, Gender == "M", select = c(Shoe_Size, Height))
print("Male data:")
print(males)
```



# 1d. Calculate means
```{r}
mean_shoe_size <- mean(data$Shoe_Size)
mean_height <- mean(data$Height)
cat("Mean Shoe Size:", mean_shoe_size, "\n")
cat("Mean Height:", mean_height, "\n")
```



# 1e. Check relationship between shoe size and height
```{r}
correlation <- cor(data$Shoe_Size, data$Height)
if (abs(correlation) < 0.1) {
cat("No significant linear relationship between shoe size and height (Correlation:", correlation, ")\n")
} else {
cat("There is a significant relationship (Correlation:", correlation, ")\n")
}
```


# 2. FACTOR
```{r}
Months <- c("March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August")
factor_Months <- factor(Months)
print("Factor Months:")
print(factor_Months)
```



# 3. Summary
```{r}
cat("Summary of Months:\n")
print(summary(Months))
cat("Summary of Factor Months:\n")
print(summary(factor_Months))
```



# 4. Vector and Frequency
```{r}
Directions <- c("East", "West", "North")
Frequency <- c(1, 4, 3)
cat("Directions:\n")
print(Directions)
cat("Frequency:\n")
print(Frequency)
```



# 4. Factor with specific order
```{r}
factor_data <- factor(Directions, levels = c("East", "West", "North"))
print("Ordered Factor Data:")
print(factor_data)

```


# 5. Read CSV data
```{r}
data <- read.table("/cloud/project/Worksheet 4/import_march (1).csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
print("Imported Data:")
print(head(data))
```



# 6. Exhaustive search function
```{r}

exhaustive_search <- function(selected_number) {
if (selected_number < 1 || selected_number > 50) {
return("The number selected is beyond the range of 1 to 50")
} else if (selected_number == 20) {
return("TRUE")
} else {
return(as.character(selected_number))
}
}

set.seed(Sys.time())
random_number <- sample(1:50, 1)
cat("The chosen number is:", random_number, "\n")
result <- exhaustive_search(random_number)
cat("Result:", result, "\n")

```


# 7. Minimum bills function
```{r}

min_bills <- function(price) {
bills <- c(1000, 500, 200, 100, 50)
bill_count <- 0

if (price %% 50 != 0) {
return("Price must be a multiple of 50.")
}

for (bill in bills) {
while (price >= bill) {
price <- price - bill
bill_count <- bill_count + 1
}
}

return(bill_count)
}

price_of_snack <- 2700
cat("Minimum number of bills needed:", min_bills(price_of_snack), "\n")

```


# 8a. Create data frame for student grades
```{r}
grades_data <- data.frame(
Name = c("Annie", "Thea", "Steve", "Hanna"),
Grade1 = c(85, 75, 75, 95),
Grade2 = c(65, 75, 55, 75),
Grade3 = c(85, 90, 80, 100),
Grade4 = c(100, 90, 85, 90)
)

print("Student Grades Data:")
print(grades_data)

```


# 8b. Average grade calculation
```{r}
cat("Students with average grade >= 88.75:\n")
avg_scores <- rowMeans(grades_data[ , 2:5])  # Calculate average scores for all students

# Loop through the students and print those with an average >= 88.75
for (i in seq_along(avg_scores)) {
  if (avg_scores[i] >= 88.75) {
    cat(grades_data$Name[i], "'s average grade this semester is", round(avg_scores[i], 2), "\n")
  }
}

```



# 8c. Find highest scores
```{r}
cat("Students with highest score > 90:\n")
results <- c()
for (i in 1:nrow(grades_data)) {
highest_score <- max(grades_data[i, 2:5])
if (highest_score > 90) {
results <- c(results, paste(grades_data$Name[i], "’s highest grade this semester is", highest_score))
}
}

cat(results, "\n")

```

#8d. Output of students' highest grade this semester
```{r}
students_scores <- data.frame(
  student = c("Annie", "Bob", "Cathy", "David"),
  semester_scores = I(list(c(88, 95, 80), c(72, 85, 91), c(90, 88, 92), c(89, 70, 78)))
)

for (i in 1:nrow(students_scores)) {
  scores <- students_scores$semester_scores[[i]]
  highest_score <- sort(scores, decreasing = TRUE)[1]  
  
  if (highest_score > 90) {
    cat(students_scores$student[i], "'s highest grade this semester is ", highest_score, ".\n", sep = "")
  }
}


```

